---
title: "Example of use of phenomenology package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example of use of phenomenology package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

First load `phenomenology` package. It provides some function to learn microeconomics using R.

```{r setup}
library(phenomenology)
```

After loading it, we must understand how it works.

1.  Consider the next equation:

    $$P = \alpha + \beta Q$$ In economics there's this convention to plot prices in the `y` axis. But we are interested in how quantities change as the prices changes. Like in this expression:

    $$Q = \frac{\alpha}{\beta } - \frac{1}{\beta }P$$ The sign of $\beta$ is the key. It determine if you are talking about supply (positive) or demand (negative).

2.  Functions inside `phenomenology` package work using the first definition. Then:

    -   $\beta$ becomes `slope` argument and it's the rate at which prices fall or increase due to changes in quantity.

    -   $\alpha$ turns in `price_q0` and it's the price that leads to a complete fall on quantity, also known as the intercept with the `y` axis.

    Having this definitions in mind let's create our first market example:

```{r demand}

demand <- create_market(price_q0 = 100,
              slope = -2)

str(demand)

```

The `create_market` function is our first step. Above I defined a demand curve following the next definition:

$$
P = 100 - 2Q_D
$$

But markets are composed by at least two parts, so we must add our supply curve:

$$
P = 30 + 3.5Q_s
$$

This is easily done with the `create_market` function:

```{r market_def}
market <- create_market(
  price_q0 = c(100,30), #
  slope = c(-2,3.5))

str(market)
```

Now we have a little more complex structure. As you can see when I declared just the demand curve, I got a list of 5 elements that defines the curve (see package documentation). Now I got a nested list with that same structure but defined for each pair of `price_q0` and `slope`. And so easy you have your market definition.

Until now we build our materials to make the analysis. Now let me introduce the `supply_and_demand` package.

```{r analysis}
analisis <- supply_and_demand(market = market,
             market_name = "Chocolate market in black market")

names(analisis)
```

This new element has some interesting outputs (check package documentation). Here we will focus in `market:`

```{r plot_market,fig.width= 7,fig.height= 5}
plot <- analisis$market

print(plot)
```
## Interpretation

Interesting result. Let's apply some concepts. The point *E1* is the equilibrium in the market. In this point consumers and suppliers have agreed in a price of `r round(dplyr::pull(analisis$cross_points,optim_p))` and the quantity demanded or supplied in that point is 
`r round(dplyr::pull(analisis$cross_points,optim_q))`
